Online Classification of arXiv Papers Based on Your Preferences

### 2014-05-16 Paper.tex
Fixed some typos and set up bibliography.

\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\pagestyle{myheadings}
\markright{Joseph McDonald}

@@

Let $I_1$ be the set of $t$ where a mistake was made and
$\widehat{y}=1$ (false positives).  Let $I_2$ be the set where a mistake was made with
$\widehat{y} = -1$ and $k = 1$ (false negatives where the algorithm updates), and $I_3$ the set where a mistake was
made with $\widehat{y} = -1$ and $k = 0$ (false negatives where the algorithm does {\it not} make an update). Finally, let $M_i = |I_i|$ for $i =
1,2,3$. We bound $M_1 + M_2$ in the same way we bound $M$ in the original Perceptron algorithm, because these are all the mistakes after which an update is made.. Assume there exists a $\rho$ and $R$ satisfying the assumptions of Theorem 7.8 from \cite{FML}, and a separating hyperplane determined by the vector $v$. Let $M' = M_1 + M_2$, then
\begin{align*}
M' \rho =&\ \frac{v}{\|v\|}\sum_{I_1 \cup I_2} y_tx_t \
 \leq   \Bigl\|\sum_{t \in I_1 \cup I_2} y_t x_t\Bigr\| \\
=& \ \Bigl\|\sum_{t \in I_1 \cup I_2} (w_{t+1} - w_t) \Bigr\| 
=  \|w_{T+1} \| \\
=&\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t+1}\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t} + y_t x_t\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} 2y_t w_t\cdot x_t - \|x_t\|^2} \\
\leq &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|x_t\|^2}
\leq \sqrt{M'r^2}.
\end{align*}
This implies that $M' \leq \frac{r^2}{\rho^2}$. Since $M = M' + M_3$, a bound on $M_3$ is needed. We revert to a weaker bound and consider the expectation of
$M_3$. Since 
\[M_3 = \sum_{t=1}^T 1_{\left\{\widehat{y}_t = -1 \cap y_t = 1\right\}}1_{k_t = 0}, \]
we get that
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 = &\ (1 - p)\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
where the last equality uses the independence of $k_t$ from the data, and the expecation is taken over the realizations of $k_t$. To control the sum of expectations in this expression we consider the expectation of $M_2$. Since
\[
M_2 =  \sum_{t=1}^T 1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}1_{k_t = 1},
\]
then
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
=&\ p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Thus
\[
E[M_3] = \ \frac{1-p}{p}\ \E[M_2].
\]
We can apply the crude bound $\E[M_3] \leq \frac{1-p}{p}\ \E[M']$, since $M_2 \leq M'$. Thus the expected number of mistakes can be bounded by
\begin{align*}
\E[M]  =  &\ \E[M'] + \E[M_3] \\
\leq & \ \frac{r^2}{\rho^2} + \frac{1-p}{p}\frac{r^2}{\rho^2} \\
\leq  & \ \frac{1}{p}\frac{r^2}{\rho^2}
\end{align*}
Note that this is a weak bound.  Not only is it greater than the
original perceptron bound by a factor of $\frac{1}{p}$, it is also only a
bound in expectation.

@@

\begin{figure}[h]\centering
\includegraphics[width=0.65\textwidth]{EveryoneRecallPaper.pdf}
\includegraphics[width=0.65\textwidth]{EveryonePrecisionPaper.pdf}
\label{fig:PrecisionandRecall}
\caption{Performance of tested online classification algorithms using different values of $p$.  The adaptive $p$ algorithm performs relatively well compared to the original algorithm ($p = 1$), and requires less than half as many negative updates. The $p = 0.1$ approach has ver poor recall, but requires far fewer updates.  The majority vote classifier combines all others in this figure.}
\end{figure}

\section{Conclusions}
To summarize our work, we're pleased to have a developed a tool that we believe can be of use for many people across several fields. We use online learning algorithms to incrementally learn the user's preferences and suggest interesting articles daily.\\

In order for this approach to be useful, it cannot present the user with every single prediction.  Instead, we weaken the algorithms to present the user with all articles it predicts will be interesting, and only some of those that it predicts will not be interesting.  We show that in this context, a bound on the expected number of mistakes can still be achieved under the same assumptions as Theorem 7.8 from \cite{FML}.  Finally, we test several algorithms on labeled data, and show that they perform comparably to the original learning algorithms, but require less input from the user.  \\
In the future, it would be nice to develop tighter theoretical bounds on our modified learning algorithm, and particularly to develop an improved bound on the adaptive $p$ algorithm.  It is possible that such a bound would lead to a more informed decision on how to change $p$ in the algorithm to minimize mistakes. Perhaps giving a different functional form to the way $p$ is changed can produce an algorithm that minimizes the number of labels required from the user, and potentially give a clear improvement over the fixed $p$ case.\\
Finally, it would be worthwhile to explore other online learning algorithms and possibly other features that might improve predictive performance. Considering a wider variety of text classification models would probably produce better results, and the limited time to develop this tool compromised our ability to develop and play around with some of these models.
\bibliography{paper}
\end{document}

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{cite}
\pagestyle{myheadings}
\markright{Steven Delong and Joseph McDonald}

@@

Let $I_1$ be the set of $t$ where a mistake was made and
$\widehat{y}=1$ (false positives).  Let $I_2$ be the set where a mistake was made with
$\widehat{y} = -1$ and $k = 1$ (false negatives where the algorithm updates), and $I_3$ the set where a mistake was
made with $\widehat{y} = -1$ and $k = 0$ (false negatives where the algorithm does {\it not} make an update). Finally, let $M_i = |I_i|$ for $i =
1,2,3$. We bound $M_1 + M_2$ in the same way we bound $M$ in the original Perceptron algorithm, because these are all the mistakes after which an update is made.. Assume there exists a $\rho$ and $R$ satisfying the assumptions of Theorem 7.8 from ~\cite{FML}, and a separating hyperplane determined by the vector $v$. Let $M' = M_1 + M_2$, then
\begin{align*}
M' \rho =&\ \frac{v}{\|v\|}\sum_{I_1 \cup I_2} y_tx_t \
 \leq   \Bigl\|\sum_{t \in I_1 \cup I_2} y_t x_t\Bigr\| \\
=& \ \Bigl\|\sum_{t \in I_1 \cup I_2} (w_{t+1} - w_t) \Bigr\| 
=  \|w_{T+1} \| \\
=&\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t+1}\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t} + y_t x_t\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} 2y_t w_t\cdot x_t - \|x_t\|^2} \\
\leq &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|x_t\|^2}
\leq \sqrt{M'r^2}.
\end{align*}
This implies that $M' \leq \frac{r^2}{\rho^2}$. Since $M = M' + M_3$, a bound on $M_3$ is needed. We revert to a weaker bound and consider the expectation of
$M_3$. Since 
\[M_3 = \sum_{t=1}^T 1_{\left\{\widehat{y}_t = -1 \cap y_t = 1\right\}}1_{k_t = 0}, \]
we get that
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 = &\ (1 - p)\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
where the last equality uses the independence of $k_t$ from the data, and the expecation is taken over the realizations of $k_t$. To control the sum of expectations in this expression we consider the expectation of $M_2$. Since
\[
M_2 =  \sum_{t=1}^T 1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}1_{k_t = 1},
\]
then
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
=&\ p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Thus
\[
E[M_3] = \ \frac{1-p}{p}\ \E[M_2].
\]
We can apply the crude bound $\E[M_3] \leq \frac{1-p}{p}\ \E[M']$, since $M_2 \leq M'$. Thus the expected number of mistakes can be bounded by
\begin{align*}
\E[M]  =  &\ \E[M'] + \E[M_3] \\
\leq & \ \frac{r^2}{\rho^2} + \frac{1-p}{p}\frac{r^2}{\rho^2} \\
\leq  & \ \frac{1}{p}\frac{r^2}{\rho^2}
\end{align*}
Note that this is a weak bound.  Not only is it greater than the
original perceptron bound by a factor of $\frac{1}{p}$, it is also only a
bound in expectation.

@@

\begin{figure}[h!]\centering
\includegraphics[width=0.65\textwidth]{EveryoneRecallPaper.pdf}
\includegraphics[width=0.65\textwidth]{EveryonePrecisionPaper.pdf}
\label{fig:PrecisionandRecall}
\caption{Performance of tested online classification algorithms using different values of $p$.  The adaptive $p$ algorithm performs relatively well compared to the original algorithm ($p = 1$), and requires less than half as many negative updates. The $p = 0.1$ approach has very poor recall, but requires far fewer updates.  The majority vote classifier combines all others in this figure.}
\end{figure}

\section{Conclusions}
To summarize our work, we're pleased to have a developed a tool that we believe can be of use for many people across several fields. We use online learning algorithms to incrementally learn the user's preferences and suggest interesting articles daily.\\
In order for this approach to be useful, it cannot present the user with every single prediction.  Instead, we weaken the algorithms to present the user with all articles it predicts will be interesting, and only some of those that it predicts will not be interesting.  We show that in this context, a bound on the expected number of mistakes can still be achieved under the same assumptions as Theorem 7.8 from ~\cite{FML}.  Finally, we test several algorithms on labeled data, and show that they perform comparably to the original learning algorithms, but require less input from the user.  \\
In the future, it would be nice to develop tighter theoretical bounds on our modified learning algorithm, and particularly to develop an improved bound on the adaptive $p$ algorithm.  It is possible that such a bound would lead to a more informed decision on how to change $p$ in the algorithm to minimize mistakes. Perhaps giving a different functional form to the way $p$ is changed can produce an algorithm that minimizes the number of labels required from the user, and potentially give a clear improvement over the fixed $p$ case.\\
Finally, it would be worthwhile to explore other online learning algorithms and possibly other features that might improve predictive performance. Considering a wider variety of text classification models would probably produce better results.
\bibliography{paper}{}
\bibliographystyle{plain}
\bibstyle{plain}
\end{document}
